Competing Bandits: The Perils of Exploration under Competition

### 2018-09-05 paper/content/barriers.tex
more small fixups
stuff

\section{Data and Reputation as Barriers to Entry}\label{section:6}

An alternative interpretation of the results in the previous section is that the temporary monopoly provides a first mover advantage to the incumbent firm. The early entry allows the incumbent to acquire data on potentially suboptimal actions without incurring an immediate reputational consequence as well as the ability to eventually build up a high reputation before the entrant enters the market. Thus, the early entry gives the incumbent both a \textit{data} advantage and a \textit{reputational} advantage over the entrant. This gives new intuition for the classic ``first mover advantage" \cite{kerin1992first} in the digital economy where data can serve as a barrier to entry in many markets.

While combined these two advantages serve as strong barriers to entry,
a natural question to ask is would having only the data advantage or
only the reputation advantage still serve as such strong barriers to
entry? Which of the two serves as a stronger barrier to entry and
under what conditions?

\swedit{To answer these questions, we isolate the effects from the two
  advantages and run two additional experiments. In the first
  experiment (called \emph{reputation erased experiment}), the
  reputation of the incumbent is artificially reset upon the entry of
  the entrant \gaedit{so that the incumbent's reputation score is not influenced by the rewards during the incumbency periods and only determined by the rewards generated in the warm start periods}. In the second experiment (called \emph{information
    erased experiment}), the information gained by the incumbent is
  erased in the sense that its posterior is reset to the initial
  (fake) prior upon the entry of the entrant.}

\swcomment{Guy, could you make it clear what the first ``erase'' mean? Like reset to default?}

\begin{table*}[h]
\centering
\begin{tabular}{|c|c|c|c||c|c|c|}
\hline
  & \multicolumn{3}{c}{Information erased}
  & \multicolumn{3}{c}{Reputation erased}\\
\hline
& $\TS$  & $\DEG$  & $\DG$ 
& $\TS$  & $\DEG$  & $\DG$
\\\hline
$\TS$ 
    & \makecell{\textbf{0.021}$\pm$0.009}
    & \makecell{\textbf{0.16}$\pm$0.02} 
    & \makecell{\textbf{0.21} $\pm$0.02} 
    %%
    & \makecell{\textbf{0.0096}$\pm$0.006}
    & \makecell{\textbf{0.11}$\pm$0.02}
    & \makecell{\textbf{0.18}$\pm$0.02} 
    \\ \hline
$\DEG$ 
    & \makecell{\textbf{0.26}$\pm$0.03} 
    & \makecell{\textbf{0.3}$\pm$0.02} 
    & \makecell{\textbf{0.26}$\pm$0.02} 
    %%
    & \makecell{\textbf{0.073}$\pm$0.01}
    & \makecell{\textbf{0.29}$\pm$0.02}
    & \makecell{\textbf{0.25}$\pm$0.02} 
    \\ \hline
$\DG$ 
    & \makecell{\textbf{0.34}$\pm$0.03} 
    & \makecell{\textbf{0.4}$\pm$0.03} 
    & \makecell{\textbf{0.33}$\pm$0.02} 
    %%
    & \makecell{\textbf{0.15}$\pm$0.02}
    & \makecell{\textbf{0.39}$\pm$0.03}
    & \makecell{\textbf{0.33}$\pm$0.02}
    \\\hline
\end{tabular}
\caption{Information-erased vs. reputation-erased experiment, on Heavy-Tail MAB instance. Each cell describes the duopoly game between the entrant's algorithm (the {\bf row}) and the incumbent's algorithm (the {\bf column}). The cell specifies the entrant's market share in this game: the average (in bold) and the 95\% confidence interval. NB: smaller average is better for the incumbent.}
\label{info_erase}
\end{table*}

@@

Tables \ref{info_erase} and \ref{rep_erase} display the results of these experiments. The reason for the finding is that, under $\TS$, the early entry allows the incumbent to do exploration without immediate reputational consequence in the early rounds and so, even when reputation is erased, the incumbent has already acquired sufficient information that simply exploiting that information allows her to quickly recover her reputation. However, while the reputational advantage gives the incumbent some cushion for doing exploration the incumbent is still vulnerable to getting unlucky while doing exploration and suffering the reputational consequences. This only happens under $\TS$ and not under $\DEG$ or $\DG$ since $\TS$, by construction, is more intelligent about the data that it acquires compared to $\DEG$ and $\DG$.

>>>>

\section{Data and Reputation as Barriers to Entry}\label{section:6}
In the temporary monopoly setting, the incumbent firm enjoys a first
mover advantage, which allows the firm to acquire data on potentially
suboptimal actions without incurring an immediate reputational
consequence, and to build up a high reputation before the entrant
enters the market. Thus, the early entry gives the incumbent both a
\textit{data} advantage and a \textit{reputational} advantage over the
entrant. This provides a quantitative perspective on the classic
``first mover advantage" \cite{kerin1992first} in the digital economy
where data can serve as a barrier to entry in many markets.

Together the two advantages serve as strong barriers to entry, and a
natural question is whether either one of the advantages alone suffice
as a strong barrier to entry. Also, which of the two is the stronger
barrier?

{To answer these questions, we isolate the effect from each of the two
  advantages and run two additional experiments. In the first
  experiment (called \emph{reputation erased experiment}), the
  reputation of the incumbent is artificially reset upon the entry of
  the entrant {so that the incumbent's reputation score is only based
    on the rewards generated in the warm start period}. In the second
  experiment (called \emph{information erased experiment}), the
  information gained by the incumbent is erased upon the entry of the
  entrant in the sense that its posterior is reset to the initial
  (fake) prior.}

\begin{table*}[h]
\centering
\begin{tabular}{|c|c|c|c||c|c|c|}
\hline
  & \multicolumn{3}{c|}{Information erased}
  & \multicolumn{3}{c|}{Reputation erased}\\
\hline
& $\TS$  & $\DEG$  & $\DG$ 
& $\TS$  & $\DEG$  & $\DG$
\\\hline
$\TS$ 
    & \makecell{\textbf{0.021}$\pm$0.009}
    & \makecell{\textbf{0.16}$\pm$0.02} 
    & \makecell{\textbf{0.21} $\pm$0.02} 
    %%
    & \makecell{\textbf{0.0096}$\pm$0.006}
    & \makecell{\textbf{0.11}$\pm$0.02}
    & \makecell{\textbf{0.18}$\pm$0.02} 
    \\ \hline
$\DEG$ 
    & \makecell{\textbf{0.26}$\pm$0.03} 
    & \makecell{\textbf{0.3}$\pm$0.02} 
    & \makecell{\textbf{0.26}$\pm$0.02} 
    %%
    & \makecell{\textbf{0.073}$\pm$0.01}
    & \makecell{\textbf{0.29}$\pm$0.02}
    & \makecell{\textbf{0.25}$\pm$0.02} 
    \\ \hline
$\DG$ 
    & \makecell{\textbf{0.34}$\pm$0.03} 
    & \makecell{\textbf{0.4}$\pm$0.03} 
    & \makecell{\textbf{0.33}$\pm$0.02} 
    %%
    & \makecell{\textbf{0.15}$\pm$0.02}
    & \makecell{\textbf{0.39}$\pm$0.03}
    & \makecell{\textbf{0.33}$\pm$0.02}
    \\\hline
\end{tabular}
\caption{Information-erased vs. reputation-erased experiment, on Heavy-Tail MAB instance. Each cell describes the duopoly game between the entrant's algorithm (the {\bf row}) and the incumbent's algorithm (the {\bf column}). The cell specifies the entrant's market share in this game: the average (in bold) and the 95\% confidence interval. NB: smaller average is better for the incumbent.}
\label{barrier_exp}
\end{table*}

@@

Table \ref{barrier_exp} display the results of these experiments. The reason for the finding is that, under $\TS$, the early entry allows the incumbent to do exploration without immediate reputational consequence in the early rounds and so, even when reputation is erased, the incumbent has already acquired sufficient information that simply exploiting that information allows her to quickly recover her reputation. However, while the reputational advantage gives the incumbent some cushion for doing exploration the incumbent is still vulnerable to getting unlucky while doing exploration and suffering the reputational consequences. This only happens under $\TS$ and not under $\DEG$ or $\DG$ since $\TS$, by construction, is more intelligent about the data that it acquires compared to $\DEG$ and $\DG$.
